### 2025-02-26 pca.qmd
20250226 - edits
20250226 - ensuring the variables are on the same scale

@@

The key distinction between [factor analysis](#sec-factorAnalysis) and PCA is depicted in @fig-pcaVsFactorAnalysis.\index{factor analysis}\index{principal component analysis}

>>>>

## Ensuring the Variables are on the Same Scale {#sec-pcaStandardize}

Before performing a PCA, it is important to ensure that the variables included in the PCA are on the same scale.
PCA seeks to identify components that explain variance in the data, so if the variables are not on the same scale, some variables may contribute considerably more variance than others.
A common way of ensuring that variables are on the same scale is to standardize them using, for example,  [*z*-scores](#sec-zScores).

@@

Both [factor analysis](#sec-factorAnalysis) and PCA can be used for data reducation.
The key distinction between [factor analysis](#sec-factorAnalysis) and PCA is depicted in @fig-pcaVsFactorAnalysis.\index{factor analysis}\index{principal component analysis}
